\chapter{Introduction} \label{chap-introduction}
This book focuses on the fundamentals of Scientific High-Performance Computing from two different 
perspectives; Computer Science and Applied Mathematics. It starts with Computer Science related 
areas such as fundamentals of High-Performance Computing, \texttt{Linux}, programming languages 
and compilers. Then, we delve into mathematical modelling and scientific applications. Finally, 
we provide an analysis of the current advances in Cloud Computing, and explain its potential 
impact on \texttt{HPC} workloads.
\newline \newline 
Each chapter is devoted to a specific field and gives detailed explanation about its subject.
Unfortunately each of these subjects are very broad to cover all of their aspects within a
chapter. Therefore, we highly suggest interested readers to refer to other references for a 
complete study. The main intention of this book is to explain how different components of an 
\texttt{HPC} system should be designed for efficient solutions to computationally demanding 
problems.
\newline \newline 
Current chapter contains a summary of each section. Picky readers can have a look at this part to
gather a general idea about a certain section if they don't want to read the whole book. However,
it is encouraged to follow the order in the book since chapters are connected to each other.
\newline \newline 
Chapter \ref{chap-hpc} begins with clarification of the fundamental concept in this book; 
High-Performance Computing which is also referred as \texttt{HPC}. Detailed answers to the
questions like "What is HPC" or "Why/When is HPC needed" can be found in this section. This
chapter also introduces the cluster concept, and gives information about hardware and software
components of HPC clusters.  
\newline \newline 
Chapter \ref{chap-linux} is about operating systems, however it specifically focuses on 
\texttt{Linux}. This chapter starts with a historical background, then briefly explains the 
evolution of \texttt{Linux}. Additionally, some basic information about kernel is given in this 
chapter. Moreover, chapter continues with the shell environment which is probably the most important 
tool that an \texttt{HPC} user should be comfortable with.
\newline \newline 
Chapter \ref{chap-programming} is an introduction to the basic programming concepts
based on commonly used languages like \texttt{C}, \texttt{C++} and \texttt{Python}. 
\newline \newline 
Chapter \ref{chap-compilers} is devoted to how compilers work. Therefore, this chapter can be thought
as a bridge between the two concepts previously discussed; operating systems and programming 
languages. Our main focus will be on the \texttt{GCC}, however other compiler suites are discussed 
as well. In additional to the compilers, we talk about debuggers and profilers too.
\newline \newline 
Chapter \ref{chap-math} is about mathematical model in science and engineering.
\newline \newline 
Chapter \ref{chap-scientific} gives information about importance of scientific libraries. We explain
commonly used libraries in this chapter.
\newline \newline 
Chapter \ref{chap-applications} explains the difference between libraries and applications. Then
we provide information about widely common scientific applications such as \texttt{OpenFOAM} 
and \texttt{CP2K}.
\newline \newline 
Chapter \ref{chap-cloud} is about cloud computing and its potential impact on \texttt{HPC}. In this
chapter, we focus on virtualization and containerization. Then we explain cloud native technologies 
like \texttt{Kubernetes} and see how it is adapted to the \texttt{HPC} workloads by major vendors.
\newline \newline 
Chapter \ref{chap-final}
